Prospects of detecting gamma-ray emission from galaxy clusters: cosmic rays and dark matter annihilations

### 2011-05-16 paper.tex
text

\subsection{Scaling relations}
First we focus on gamma-ray flux-cluster mass scaling relations for DM
annihilation and CR induced emission in
Fig.~\ref{fig:lum_mass_scaling}.  
\begin{figure*}
  \includegraphics[width=0.99\columnwidth]{figures/MLscaling.100M.eps}
  \includegraphics[width=0.99\columnwidth]{figures/MLscaling.10G.eps}
  \caption{\it Scaling relations of the cluster's virial mass $\mvir$ and the
    gamma-ray luminosity for energies $E_\gamma>100$~MeV (left) and
    $E_\gamma>10$~GeV (right).  Shown are the relations for the CR-induced
    emission that is dominated by pion decay resulting from hadronic CR
    interactions (solid), the leptophilic model of DM annihilations that
    includes the scaling of the substructure and Sommerfeld boost factors with
    cluster mass from Eq.~(\ref{eq:LP_scaling}) (dash-dotted) as well as the
    benchmark model $K'$ of DM annihilation that includes the scaling of the
    substructure boost with cluster mass of Eq.~(\ref{eq:BM_scaling})
    (dashed). Note that the scaling of this particular benchmark model is
    indicative for all (non-Sommerfeld enhanced) DM annihilation models while
    the normalization depends on the particular cross section and neutralino
    mass.}
\label{fig:lum_mass_scaling}
\end{figure*}
The mass scaling of the substructure boost steepens the intrinsically
shallower DM annihilation relation of BM models
(Eq.~\ref{eq:BM_scaling}) while this steepening is in fact
over-compensated by the inverse mass scaling of the Sommerfeld boost
in LP models that prefers smaller halos with a larger ratio of
$c/v_\mathrm{circ}$ (Eq.~\ref{eq:LP_scaling}). In contrast, the CR
scaling relation is considerably steeper as shown by reference
\cite{2010MNRAS.409..449P}:
\begin{eqnarray}
L_{\gamma}(>100\,\mev) = 1.8\times10^{45}
\left(\frac{\mvir}{10^{15}\,\msun}\right)^{1.46}\mbox{erg~s}^{-1}\nonumber\\
L_{\gamma}(>10\,\gev)  = 1.4\times10^{43}
\left(\frac{\mvir}{10^{15}\,\msun}\right)^{1.34}\mbox{erg~s}^{-1}\,.\nonumber\\
\end{eqnarray}
Note that the CR luminosity scaling relations \footnote{The scaling
  relations show the flux inside $\rvir$ and do not include the
  contribution from galaxies. Also note that in this paper the scaling
  relations are normalized with $10^{15}/h_{70}\,\msun$, in contrast
  to $10^{15}/h\,\msun$ that was used in \cite{2010MNRAS.409..449P}.}
include the IC gamma-ray contribution from shock accelerated electrons
as well as secondary electrons created through CR-proton
interactions. The gamma-ray flux from the electrons is, however,
dominated by the decaying $\pi^0$ and only show up as a small
flattening of the scaling relation at energies around
$10-100\,\gev$. The steeper mass scaling of the CR compared to the DM
implies already a general strategy to minimize the CR-induced
foreground for DM annihilation and argues for very nearby groups. We
note that these simulations do not include AGN feedback that is
thought to furthermore reduce the baryon fraction in groups relative
to that in clusters \cite{2008ApJ...687L..53P}. The resulting smaller
target density for hadronic CR interactions will furthermore steepen
the $L_\gamma-M$ relation of CR induced emission, making the case for
groups even stronger.

>>>>

\subsection{Scaling relations}
First we focus on gamma-ray flux-cluster mass scaling relations for DM
annihilation and CR induced emission in
Fig.~\ref{fig:lum_mass_scaling}.  
\begin{figure*}
  \includegraphics[width=0.99\columnwidth]{figures/MLscaling.100M.eps}
  \includegraphics[width=0.99\columnwidth]{figures/MLscaling.10G.eps}
  \caption{\it Scaling relations of the cluster's virial mass $\mvir$ and the
    gamma-ray luminosity for energies $E_\gamma>100$~MeV (left) and
    $E_\gamma>10$~GeV (right).  Shown are the relations for the CR-induced
    emission that is dominated by pion decay resulting from hadronic CR
    interactions (solid), the leptophilic model of DM annihilations that
    includes the scaling of the substructure and Sommerfeld boost factors with
    cluster mass from Eq.~(\ref{eq:LP_scaling}) (dash-dotted) as well as the
    benchmark model $K'$ of DM annihilation that includes the scaling of the
    substructure boost with cluster mass of Eq.~(\ref{eq:BM_scaling})
    (dashed). Note that the scaling of this particular benchmark model is
    indicative for all (non-Sommerfeld enhanced) DM annihilation models while
    the normalization depends on the particular cross section and neutralino
    mass.}
\label{fig:lum_mass_scaling}
\end{figure*}
The mass scaling of the substructure boost steepens the intrinsically
shallower DM annihilation relation of BM models
(Eq.~\ref{eq:BM_scaling}) while this steepening is in fact
over-compensated by the inverse mass scaling of the Sommerfeld boost
in LP models that prefers smaller halos with a larger ratio of
\C{CHRISTOPH, Should it not be $\sigv$ here instead of
  $v_\mathrm{circ}$?}  $c/v_\mathrm{circ}$
(Eq.~\ref{eq:LP_scaling}). In contrast, the CR scaling relation is
considerably steeper as shown by reference \cite{2010MNRAS.409..449P}:
\begin{eqnarray}
L_{\gamma}(>100\,\mev) = 1.8\times10^{45}
\left(\frac{\mvir}{10^{15}\,\msun}\right)^{1.46}\mbox{erg~s}^{-1}\nonumber\\
L_{\gamma}(>10\,\gev)  = 1.4\times10^{43}
\left(\frac{\mvir}{10^{15}\,\msun}\right)^{1.34}\mbox{erg~s}^{-1}\,.\nonumber\\
\end{eqnarray}
Note that the CR luminosity scaling relations \footnote{The scaling
  relations show the flux inside $\rvir$ and do not include the
  contribution from galaxies. Also note that in this paper the scaling
  relations are normalized with $10^{15}/h_{70}\,\msun$, in contrast
  to $10^{15}/h\,\msun$ that was used in \cite{2010MNRAS.409..449P}.}
include the IC gamma-ray contribution from shock accelerated electrons
as well as secondary electrons created through CR-proton
interactions. The gamma-ray flux from the electrons is, however,
dominated by the decaying $\pi^0$ and only show up as a small
flattening of the scaling relation at energies around
$10-100\,\gev$. The steeper mass scaling of the CR compared to the DM
implies already a general strategy to minimize the CR-induced
foreground for DM annihilation and argues for very nearby groups. We
note that these simulations do not include AGN feedback that is
thought to furthermore reduce the baryon fraction in groups relative
to that in clusters \cite{2008ApJ...687L..53P}. The resulting smaller
target density for hadronic CR interactions will furthermore steepen
the $L_\gamma-M$ relation of CR induced emission, making the case for
groups even stronger.
